---
title: "Laslett Transform Examples"
author: "Kassel Hingee"
date: "18 January 2016"
output: html_document
---

Examples of Lastlett's Tranform applied to various RACS



```{r}
library(spatstat)
data(heather)
xi <- heather$coarse
source("laslettTransform.R")
ltxi <- laslettTransform(xi)
```

The result is a list of a `ppp` object and the transformed image before it was converted into ppp. See below plots. In the image plot (right) the tangent points in the image have value of `2`, the background has value of `0` and outside the window is `NA`. For confirmation the points in `ltxi$points` have been plotted as circles on top of the image plot.  

```{r, echo=FALSE}
layout(matrix(c(1,2),ncol=2,nrow=1))
plot(ltxi$points,cols="red",chars="+")
plot(ltxi$image)

plot(ltxi$points,add=TRUE)
```

Rotate input `xi` to have a better chance of interpreting the point pattern.

```{r}
xir <- rotate.owin(xi)
ltxir <- laslettTransform(xir)
```
```{r, echo=FALSE}
plot.new()
plot(xir)
tangentpointslist <- findlowlefttangentpts(xir)
points(tangentpointslist$X,tangentpointslist$Y,col="red",pch="+")
plot(ltxir$points,cols="red",chars="+")
```


